---
title: "PA2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


reading the dataset from web and test it
```{r }
library(ggplot2)
setwd("C://Users//pg000//Desktop//Download")
fileUrl <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
destfile <- ".//repdata-data-StormData.csv.bz2"
if(!file.exists(destfile)) {
 download.file(fileUrl, destfile = destfile, quiet = TRUE)
 dateDownload <- date()
}
rawData <- read.csv(bzfile(destfile), stringsAsFactors = FALSE)
names(rawData)
```



```{r }
rawData$Total_cas <- rawData$FATALITIES + rawData$INJURIES
value <- function(x) {
 x <- tolower(x)
 if(x=="k")res <- 1000
 if(x == "m")res <- 1e+06
 if(x == "b") res <- 1e+09
 else res <- 1
 res
}
rawData$pd <- rawData$PROPDMG * sapply(rawData$PROPDMGEXP, value)/1000000
rawData$cd <- rawData$CROPDMG * sapply(rawData$CROPDMGEXP, value)/1000000
rawData$td <- rawData$pd + rawData$cd
```
## taking relevant variable new dataset is constructed 
```{r}
proc_data <- rawData[,c("EVTYPE","FATALITIES","INJURIES",
                        "Total_cas","pd","cd","td")]
proc_data <- aggregate(proc_data[,2:7], 
           by =list(proc_data$EVTYPE), FUN = sum, na.rm = TRUE)

colnames(proc_data) <- c("EVETYPE", colnames(proc_data[2:7]))
```
top_data() function takes data frame (df), column number (col) and returns the top results.

```{r}
top_data <- function(df, col,top) {
 df <- df[,c(1,col)]
 df <- df[order(df[,2],decreasing = TRUE),]
 df <- df[1:top,]
 rownames(df) <- NULL
 df
}
```
## top 3 events with FATALITIES
```{r}
top_data(proc_data,2,3)
```
## check the INJURIES
```{r}
top_data(proc_data,3,3)
```
## top 10 events with total casualities
```{r}
cu_data <- top_data(proc_data, 4, 10)
cu_data
```
## plot the top 10 events with most total casualities
```{r}
ggplot(cu_data, aes(x=EVETYPE, y=Total_cas, fill = EVETYPE))+
geom_bar(stat = "identity")+
ggtitle("Top due to sever weather events in US")+
 xlab("Events")+
 ylab("Total Casulties")+
 theme(axis.text.x = element_text(angle = 90, vjust =0.5))
```
## show the United States which the types of events have greatest consequence 
```{r}
top_data(proc_data, 5,3)
```
## the top 10 event with most total_cas
```{r}
top_data(proc_data, 6,3)
dmg_data <- top_data(proc_data, 7,10)
dmg_data
```
## plot the graph

```{r}
ggplot(dmg_data, aes(x = EVETYPE, y = td, fill =EVETYPE))+
 geom_bar(stat ="identity") +
 ggtitle("weather economic consequence events in US") +
 xlab("events")+ 
 ylab("total damage in million USD")+
 theme(axis.text.x = element_text(angle = 90, vjust =0.5))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
